\chapter{Mechanics}

\section{Miscellaneous}

\[
	\text{\% error} = \frac{\text{observed} - \text{theoretical}}{\text{theoretical}} * 100\%
\]

\section{Kinematics}

\[
	x(t) = \frac{a}{2}t^2 + v_0 t + x_0 \qquad
	v(t) = v_0 + at
\]\[
	v^2 = v_0^2 + 2a\Delta x \qquad
	\Delta x = \frac{v_0 + v}{2} * \Delta t
\]

\section{Forces}

\[
	F_\text{net} = ma
\]

$F_T$ represents tension. It always points in the direction on which the rope pulls on the object.

$F_N$ represents normal force. It takes the direction and magnitude necessary to prevent the object from passing through the surface that creates the normal force.

\subsection{Friction}

\[
	f_s \leq \mu_s F_N
\]

The static friction takes the direction and magnitude necessary to prevent the object from moving in the component parallel to the surface, until the magnitude reaches $\mu_s F_N$. Upon reaching $\mu_s F_N$, the static friction is replaced by kinetic friction and the object starts moving:

\[
	k_s = \mu_k F_N
\]

\subsection{Centripetal force}

The centripetal force always points towards the center of the circle representing the object's path, and therefore is perpendicular to the velocity. It is just another name for the net force in the centripetal direction. The centripetal acceleration is what causes the object to rotate.

\[
	F_c = ma_c = m \frac{v^2}{r} = mr\omega^2 \qquad a_c = \frac{v^2}{r}
\]

\section{Work and Energy}

\[
	W = \int_a^b F(\vec{r}) \cdot d\vec{r}
\]

In one dimension:

\[
	W = \int_a^b F(x) dx
\]

\begin{namedtheorem}[Work-energy theorem] The work is the change in kinetic energy: 
	\[
		W = \Delta K
	\]
\end{namedtheorem}

\section{Simple Harmonic Motion}

The object is at rest at the \textbf{equilibrium position}. $x = 0$ when the object is at equilibrium, and $x$ represents the distance and direction from the equilibrium. When you pull it to one direction, the \textbf{restoring force} pulls the object back toward the equilibrium position. It oscillates back and forth, between $x = -A$ and $x = A$, where $A$ is the \textbf{amplitude}. The \textbf{period} is the amount of time to complete one oscillation, and the \textbf{frequency} is the amount of oscillations that happen in one second (or some other time unit).

\begin{align*}
	x(t) &= A \cos(\omega t + \phi) \\
	v(t) &= -A\omega \sin(\omega t + \phi) \\
	a(t) &= -A \omega^2 \cos(\omega t + \phi)
\end{align*}

\subsection{Spring force}

\begin{align*}
	F_s &= -kx \\
	U_s &= -\tfrac{1}{2} kx^2 \\
\end{align*}

\begin{align*}
	a &= - \frac{k}{m} x \\
	v &= \sqrt{\frac{k}{m} (A^2 - x^2)} \\
\end{align*}

\section{Simple Harmonic Motion (old)}

\[
	x = A \cos(\omega t + \varphi) \quad
	v = -\omega A \cos(\omega t + \varphi) \quad
	a = -\omega^2 A \cos(\omega t + \varphi)
\]
\[
	x_{\max} = A \qquad
	v_{\max} = \omega A \qquad
	a_{\max} = \omega^2 A \qquad
	F_{\max} = m\omega^2 A
\]

\subsection{Springs and Slinkies}

$x$ represents the distance from the equilibrium.

If you put a mass on top of the slinky, $\Delta x_\text{eq}$ represents the difference between the original equilibrium and the new equilibrium.

\[
	F_s = kx = ma \qquad
	F_{s_{\max}} = k\Delta x_\text{eq} = 9.8 \Delta m
\]
\[
	f = \frac{1}{2\pi}\sqrt{\frac{k}{m}} \qquad 
	T = 2\pi\sqrt{\frac{m}{k}} \qquad 
	\omega = 2\pi f = \sqrt{\frac{k}{m}}
\]
\[
	SPE = \frac{1}{2} kx^2 \qquad
	KE = \frac{1}{2} mv^2
\]
\[
	TME = \frac{1}{2} kx^2 + \frac{1}{2} mv^2 = \frac{1}{2} kA^2 = \frac{1}{2} mv_{\max}^2
\]

\subsection{Springs in parallel and series}

% Converted from https://en.wikipedia.org/wiki/Series_and_parallel_springs using https://mediawiki2latex.wmflabs.org/

\begin{tabular}{|>{\RaggedRight}p{0.38225\linewidth}|>{\RaggedRight}p{0.23286\linewidth}|>{\RaggedRight}p{0.23286\linewidth}|} \hline 
	{\bfseries \hspace*{0pt}\ignorespaces{}\hspace*{0pt}Quantity}&{\bfseries \hspace*{0pt}\ignorespaces{}\hspace*{0pt}In Series}&{\bfseries \hspace*{0pt}\ignorespaces{}\hspace*{0pt}In Parallel} %\endhead  
	% \\ \hline \multicolumn{4}{|>{\RaggedRight}p{0.97143\linewidth}|}{\hspace*{0pt}\ignorespaces{}\hspace*{0pt}}
	\\ \hline \hspace*{0pt}\ignorespaces{}\hspace*{0pt}Equivalent spring constant&\hspace*{0pt}\ignorespaces{}\hspace*{0pt}{${\displaystyle {\frac {1}{k_{\mathrm {eq} }}}={\frac {1}{k_{1}}}+{\frac {1}{k_{2}}}}$}&\hspace*{0pt}\ignorespaces{}\hspace*{0pt}{${\displaystyle k_{\mathrm {eq} }=k_{1}+k_{2}}$}
	% \\ \hline \hspace*{0pt}\ignorespaces{}\hspace*{0pt}Equivalent compliance&\hspace*{0pt}\ignorespaces{}\hspace*{0pt}{${\displaystyle c_{\mathrm {eq} }=c_{1}+c_{2}}$}&\hspace*{0pt}\ignorespaces{}\hspace*{0pt}{${\displaystyle {\frac {1}{c_{\mathrm {eq} }}}={\frac {1}{c_{1}}}+{\frac {1}{c_{2}}}}$}
	\\ \hline \hspace*{0pt}\ignorespaces{}\hspace*{0pt}Deflection (elongation)&\hspace*{0pt}\ignorespaces{}\hspace*{0pt}{\itshape {${\displaystyle x_{\mathrm {eq} }=x_{1}+x_{2}}$}}&\hspace*{0pt}\ignorespaces{}\hspace*{0pt}{\itshape {${\displaystyle x_{\mathrm {eq} }=x_{1}=x_{2}}$}}
	\\ \hline \hspace*{0pt}\ignorespaces{}\hspace*{0pt}Force&\hspace*{0pt}\ignorespaces{}\hspace*{0pt}{${\displaystyle F_{\mathrm {eq} }=F_{1}=F_{2}}$}&\hspace*{0pt}\ignorespaces{}\hspace*{0pt}{${\displaystyle F_{\mathrm {eq} }=F_{1}+F_{2}}$}
	\\ \hline \hspace*{0pt}\ignorespaces{}\hspace*{0pt}Stored energy&\hspace*{0pt}\ignorespaces{}\hspace*{0pt}{${\displaystyle E_{\mathrm {eq} }=E_{1}+E_{2}}$}&\hspace*{0pt}\ignorespaces{}\hspace*{0pt}{${\displaystyle E_{\mathrm {eq} }=E_{1}+E_{2}}$}\\ \hline 
\end{tabular}

\section{Center of mass}

The center of mass of an object which consists of masses at discrete points $\vec{r}_1, \vec{r}_2, \vec{r}_3, ..., \vec{r}_n$ is
\[
	\vec{r}_\text{cm} = \frac{\sum m_i \vec{r}_i}{\sum m_i}
\]

The center of mass of a solid object with variable density is
\[
	\vec{r}_\text{cm} = \frac{\int \vec{r} dm}{\int dm}
\]

To find the velocity of the center of mass, replace $\vec{r}_i$ with $\vec{v}_i$ in the above equations.

\subsection{Density}

\begin{align*}
	\text{3D: density} \qquad & \rho = \frac{M}{v} = \frac{dm}{dV} & (\frac{\kilogram}{\meter^3}) \\
	\text{2D: surface density} \qquad & \sigma = \frac{M}{A} = \frac{dm}{dA} & (\frac{\kilogram}{\meter^2}) \\
	\text{1D: linear density} \qquad & \lambda = \frac{M}{L} = \frac{dm}{dx} & (\frac{\kilogram}{\meter}) \\
\end{align*}

\section{Momentum/Impulse}

\begin{definition}[Momentum] \textbf{Momentum} ($\vec{p}$) is defined as:
	\[
		\vec{p} = m \vec{v} \qquad (\Newton \second = \kilogram \frac{\meter}{\second})
	\]
\end{definition}

\begin{definition}[Impulse] \textbf{Impulse} is the integral of force over time. By the Impulse-Momentum Theorem, it is also the change in momentum.
	\[
		\vec{J} = \int \vec{F} dt = \Delta p = \vec{F}_\text{net,avg} \Delta t
	\]
	Additionally,
	\[
		\vec{F}_\text{net} = \frac{dp}{dt}
	\]
\end{definition}

\begin{theorem}[Conservation of momentum]
	Momentum is conserved when there are no external forces acting upon the object.
\end{theorem}

\section{Rotational motion}

\subsection{Rotational kinematics}

$\vec{\theta}$, angle, is the rotational equivalent of position.

$\Delta \theta$ is called angular displacement.

\begin{align*}
\text{Angular velocity} = \vec{\omega} &= \frac{d\vec{\theta}}{dt} \quad \left(\frac{\radian}{\second}\right) \\
\text{Angular acceleration} = \vec{\alpha} &= \frac{d\vec{\omega}}{dt} \quad \left(\frac{\radian}{\second^2}\right) \\
\text{Arc length} = s &= R \theta \\
\text{Velocity} = v &= R \omega \\
\text{Acceleration} = a &= R \alpha \\
\end{align*}

\subsubsection{Constant acceleration}

\begin{align*}
	\omega_f &= \omega_0 + \alpha t \\
	\Delta \theta &= \omega_0 t + \tfrac{1}{2} \alpha t^2 \\
	\omega_f^2& = \omega_0^2 + 2 \alpha \delta \theta
\end{align*}

\subsection{Torque}

\[
	\vec{\tau} = \vec{r} \crossp \vec{F} \qquad |\tau| = |r| |F| \sin \theta
\]

\subsection{Rotational inertia}

Where $r$ represents the distance between the axis of rotation and the object, $m$ represents mass, $I$ represents rotational inertia, and $K$ represents rotational kinetic energy:

\[
	I = mr^2 \qquad I_\text{total} := \sum m_i r_i^2 = \int r^2 dm \qquad K = \frac{1}{2} I \omega^2
\]

\begin{namedtheorem}[Parallel Axis Theorem]
	If $d$ is the distance between the axis through the center of mass and the desired axis of rotation parallel to the axis through the center of mass, then the rotational inertia through the desired axis of rotation is
	\[
		I_{\text{II}} = I_{\text{cm}} + Md^2
	\]
\end{namedtheorem}

\subsection{Newton's Second Law}

\[
	\sum \tau = I \alpha
\]

\subsection{Rotational kinetic energy}

\begin{definition}[Rotational kinetic energy]
	\[
		K_{\text{rot}} = \frac{1}{2} I \omega^2
	\]
\end{definition}

\begin{definition}[Rotational work]
	\[
		W = \int \tau \,d\theta = \Delta K
	\]
\end{definition}

\subsection{Rolling}

During rolling, the velocity of the center of mass is
\[
	v_\text{cm} = R\omega
\]
in the direction of rolling. Add $v_\text{cm} = R\omega$ to the tangential velocity, $R\omega$ in the direction perpendicular to the radius.

When rolling at the maximum speed possible without slipping, the friction is
\[
	f_s = \mu_s F_N
\]
Otherwise,
\[
	f_s < \mu_s F_N
\]

\subsection{Angular momentum}

\[
	\vec{L} = I \vec{\omega} \text{ for solid objects}
\]

\[
	\vec{L} = \vec{r} \crossp \vec{p} \text{ for point masses}
\]

\[
	L = rp\sin(\theta) = rmv\sin(\theta) \text{ for point masses}
\]

Note: angular velocity's direction is specified by right-hand rule.

The derivative of angular momentum is the change in external net torque.
\[
	\frac{d\vec{L}}{dt} = \vec{\tau}_\text{net, ext}
\]

\subsection{Equilibrium}

If $\sum F = 0$ and $\sum \tau = 0$, then the object is in equilibrium and therefore is not accelerating.

If $\sum F = 0$ and $\sum \tau \neq 0$, then the object is accelerating in its rotation but has constant velocity.

If $\sum F \neq 0$ and $\sum \tau = 0$, then the object is linearly accelerating but is not accelerating rotationally.

If $\sum F \neq 0$ and $\sum \tau \neq 0$, then the object is in complex motion.

\subsection{Gravitation}

Using the equations which are similar to electrostatics:

\begin{definition}[Gravitational constant]
	\[
		G \approx 6.67 * 10^{-11} \frac{\Newton \meter^2}{\kilogram^2}
	\]
\end{definition}

\begin{definition}[Gravitational field]
  The vector field $g$, which applies to all objects.

  The electric field points towards masses.

	$g$ is also the acceleration caused by gravity.
\end{definition}

\begin{namedlaw}[Newton's law of universal gravitation]
  If stationary point or spherical masses $m_1$ and $m_2$ are near each other but not touching, then the magnitude of the gravitational force between the two objects is
  \[
    |F_g| = \frac{G m_1 m_2}{r^2}
  \]
\end{namedlaw}

\begin{law}
  The gravitational force on a stationary point or spherical mass $g$ is $F_g = mg$, where $g$ is the strength and direction of the gravitational field.
\end{law}

\begin{law}[by Newton's law of universal gravitation]
  The gravitational field at distance $r$ due to a stationary point or spherical charge $m$ is
  \[
    g = \frac{GM}{r^2}.
  \]

  The gravitational field at distance $r$ due to a massive object whose total mass is $\int dm$ is
  \[
    g = \int \frac{k \,dm}{r^2}
  \]
  where $r$'s tail is at the charge and points towards the point for which we want to calculate $g$.
\end{law}

\begin{namedtheorem}[Gauss's Law]
  The net flux is proportional to the enclosed mass, and can be equated to the flux through the surface:
  \[
    \Phi_{\text{net}} = - 4\pi G M = \oint \vec{g} \dotp d\vec{A}
  \]
\end{namedtheorem}
